---
title: "External validation against PopED and PFIM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{External validation against PopED and PFIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

LibeRality includes an executable external-validation suite. It evaluates
matched population PK designs in LibeRality, PopED and PFIM and compares the
complete Fisher information matrices rather than only their determinants.

## Harmonised convention

The comparison uses fixed effects on their natural scale, OMEGA variances and
residual variances. PFIM expresses residual-error parameters as standard
deviations; its FIM is transformed with the corresponding Jacobian before any
comparison. The `fo_block` LibeRality approximation implements the conventional
block-diagonal population-FO form used by PopED and PFIM. LibeRality's default
`full_gaussian` form remains unchanged.

## Reproduce the validation

From the ecosystem repository root:

```{r, eval=FALSE}
source("validation/liberality/external/install-dependencies.R")
source("validation/liberality/external/run-validation.R")
```

Or run it directly from R:

```{r, eval=FALSE}
result <- lity_external_validate(
  repetitions = 10,
  output_dir = "validation/liberality/external/results/manual"
)
print(result)
```

Each run writes complete matrices, numerical deltas, RSE differences, cold and
warm timings, a D-optimal grid-search result, a JSON manifest and an HTML
report. The command-line runner exits with a failure status if any declared
tolerance is exceeded.

The runtime table distinguishes setup, first (cold) evaluation, repeated warm
core evaluation, and repeated warm setup-plus-evaluation. Timings use a
high-resolution wall clock. They are useful for detecting regressions on the
same machine, but should not be interpreted as universal package rankings.

## Scope limitation

PFIM 7.0.3's `Combined1` residual model uses `(a + b*f)^2`. LibeRality and
PopED's independent combined error uses `a^2 + b^2*f^2`. The combined-error
fixture is therefore compared between LibeRality and PopED and explicitly
marked unsupported for PFIM. This is a model-convention limitation, not a
numerical discrepancy.
